---
title: "Reproducible Research Assignment Week 2"
author: "Constantin Hartmann"
date: "27 2 2020"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

1. Code for reading in the dataset and/or processing the data

Having loaded all the packages needed, I use the read.csv-function to read in the data and the ymd-function from the lubridate package to convert the date column into the appropriate date format. 

```{r echo = TRUE, results = 'markup'}
library(lubridate)
library(tidyverse)
library(dplyr)

data <- read.csv("C:\\Users\\USER\\Dropbox\\Statistik\\activity.csv", header = TRUE)
data$date <- ymd(data$date)
```

2. Histogram of the total number of steps taken each day

The distribution of the number of steps per day is as follows:

```{r echo = FALSE}

steps_per_day <- data %>%
        group_by(date) %>%
        summarize("Steps_per_Day" := sum(steps, na.rm = TRUE))
ggplot(steps_per_day, aes(x = Steps_per_Day)) + geom_histogram(binwidth = 3000) + ggtitle("Total number of steps taken each day")

```

3. Mean and median number of steps taken each day

I used the following code to compute and print the mean and median number of steps taken each day: 

```{r echo = TRUE, results = 'hide'}

steps <- data %>%
        group_by(date) %>%
        mutate(steps_per_day = sum(steps, na.rm = TRUE))

mean(steps$steps_per_day)
median(steps$steps_per_day)

```

4. Time series plot of the average number of steps taken

I combined geom_line and geom_point plots to build a time series plot of the average number of steps taken:

```{r echo = TRUE, results = 'hide'}

steps_avg <- data %>%
        group_by(date) %>%
        summarize(avg = mean(steps, na.rm = TRUE))
steps_avg        
ggplot(na.omit(steps_avg), aes(x = date, y = avg)) + 
        geom_line(color = "steelblue") + geom_point()

```

5. The 5-minute interval that, on average, contains the maximum number of steps

I wrote the following code to compute and print out the interval that, on average, contains the maximum number of steps. The respective interval is printed below:

```{r echo = TRUE}

interval_max <- data %>%
        group_by(interval) %>%
        summarize("Interval_Steps_Avg" := mean(steps, na.rm = TRUE)) %>%
        filter(Interval_Steps_Avg == max(Interval_Steps_Avg))

interval_max
```

6. Code to describe and show a strategy for imputing missing data

There are a number of ways of dealing with missing values. In this case, I chose to skip them by setting the na.rm argument to TRUE. Replacing NAs by 0 wouldn`t change anything, as neither 0 nor NA (removed) has any influence on the mean. 

```{r eval = FALSE}

interval_max <- data %>%
        group_by(interval) %>%
        summarize("Interval_Steps_Avg" := mean(steps, na.rm = TRUE)) %>%
        filter(Interval_Steps_Avg == max(Interval_Steps_Avg))

```

7. Histogram of the total number of steps taken each day after missing values are imputed

```{r echo = TRUE, results = 'hide'}
steps_avg <- data %>%
        group_by(date) %>%
        summarize("Steps_per_Day" := sum(steps, na.rum = TRUE))
```
```{r echo = FALSE}
ggplot(steps_avg, aes(x=Steps_per_Day)) + geom_histogram(binwidth = 3000)

```

8. Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r echo = TRUE, results = 'hide'}

z <- data %>%
        mutate(weekday = weekdays(date)) %>%
        mutate(daytype = case_when((weekday == "Samstag" | weekday == "Sonntag") ~ "Weekend", 
                (weekday == "Montag" | weekday == "Dienstag" | weekday == "Mittwoch" | weekday == "Donnerstag" | weekday ==                  "Freitag") ~ "Workingday"))
  
t <- z %>%
        group_by(interval, daytype) %>%
        #group_by(interval) %>%
        mutate("mean_steps" = mean(steps, na.rm=TRUE))

z$daytype <- as.factor(z$daytype)
```
```{r echo = FALSE}
ggplot(t, aes(x=interval, y=mean_steps)) + geom_line() + facet_wrap( ~ daytype)

```
9. All the R code needed to reproduce the results (numbers, plots, etc.) in the report

Finally, the complet R code I wrote:

```{r eval = FALSE}
library(lubridate)
library(tidyverse)
library(dplyr)

data <- read.csv("C:\\Users\\USER\\Dropbox\\Statistik\\activity.csv", header = TRUE)
data$date <- ymd(data$date)

steps_per_day <- data %>%
        group_by(date) %>%
        summarize("Steps_per_Day" := sum(steps, na.rm = TRUE))
ggplot(steps_per_day, aes(x = Steps_per_Day)) + geom_histogram(binwidth = 3000) + ggtitle("Total number of steps taken each day")

steps <- data %>%
        group_by(date) %>%
        mutate(steps_per_day = sum(steps, na.rm = TRUE))

mean(steps$steps_per_day)
median(steps$steps_per_day)

steps_avg <- data %>%
        group_by(date) %>%
        summarize(avg = mean(steps, na.rm = TRUE))
steps_avg        
ggplot(na.omit(steps_avg), aes(x = date, y = avg)) + 
        geom_line(color = "steelblue") + geom_point()

interval_max <- data %>%
        group_by(interval) %>%
        summarize("Interval_Steps_Avg" := mean(steps, na.rm = TRUE)) %>%
        filter(Interval_Steps_Avg == max(Interval_Steps_Avg))

interval_max

interval_max <- data %>%
        group_by(interval) %>%
        summarize("Interval_Steps_Avg" := mean(steps, na.rm = TRUE)) %>%
        filter(Interval_Steps_Avg == max(Interval_Steps_Avg))

steps_avg <- data %>%
        group_by(date) %>%
        summarize("Steps_per_Day" := sum(steps, na.rum = TRUE))

ggplot(steps_avg, aes(x=Steps_per_Day)) + geom_histogram(binwidth = 3000)

z <- data %>%
        mutate(weekday = weekdays(date)) %>%
        mutate(daytype = case_when((weekday == "Samstag" | weekday == "Sonntag") ~ "Weekend", 
                (weekday == "Montag" | weekday == "Dienstag" | weekday == "Mittwoch" | weekday == "Donnerstag" | weekday ==                  "Freitag") ~ "Workingday"))
  
t <- z %>%
        group_by(interval, daytype) %>%
        #group_by(interval) %>%
        mutate("mean_steps" = mean(steps, na.rm=TRUE))

z$daytype <- as.factor(z$daytype)

ggplot(t, aes(x=interval, y=mean_steps)) + geom_line() + facet_wrap( ~ daytype)

```

